---
title: "Getting started with surveywts"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(surveywts)
```

This article is for an R-fluent analyst who is not a survey
methodologist. It explains the objects the package returns, the two
standard workflows, how to choose among sibling methods, and the
recurring terms. The README shows the calls; this article explains what
comes back and why.

## The class system

Every surveywts function takes a survey object from
[surveycore](https://jdenn0514.github.io/surveycore/) and returns one.
Three classes cover everything:

| Class | What it is |
|---|---|
| `survey_taylor` | A probability design that computes variance by Taylor linearization (see the [glossary](#taylor-linearization)). |
| `survey_replicate` | A probability design that carries replicate weight columns (`rep_1`, `rep_2`, ...) and computes variance from their spread. |
| `survey_nonprob` | A non-probability sample with estimated weights. Replicate methods add `repwt_*` columns in place; the class never changes. |

The diagram shows the object progression for the two sample types:

```
probability sample              non-probability sample

survey_taylor                   survey_nonprob
    │  create_*_weights()           │  create_*_weights()
    ▼                               │  (quasi-randomization bootstrap,
survey_replicate                    │   grouped jackknife / DAGJK)
    │  as_taylor_design()           ▼
    ▼                           survey_nonprob + repwt_* columns
survey_taylor                       (class unchanged; variance-ready)
```

Weighting functions preserve the input class. `calibrate()`,
`adjust_nonresponse()`, `trim_weights()`, and their siblings return the
same class they receive. Only the probability-path replicate creators
change the class: they turn a `survey_taylor` into a `survey_replicate`.
`as_taylor_design()` converts a `survey_replicate` back, and it always
warns that the replicate-specific variance detail is lost in the
conversion.

One crossover exists, and it is a trap rather than a route.
`create_bootstrap_weights()` with a probability-style `type` (the
default included) accepts a `survey_nonprob`, silently wraps it as a
simple-random-sample design, and returns a `survey_replicate`. Those
replicates ignore the propensity-estimation step, and
`as_taylor_design()` refuses the result. On a non-probability sample,
use `type = "quasi-randomization"` instead — see the note in
`create_bootstrap_weights()`. Of the other creators,
`create_brr_weights()`, `create_gen_boot_weights()`,
`create_gen_rep_weights()`, and `create_sdr_weights()` reject a
`survey_nonprob` outright; `create_jackknife_weights()` rejects it for
`type = "jk1"` and `"jkn"` and accepts it only with `type = "grouped"`,
the DAGJK path the diagram shows.

## The two standard workflows

Both workflows end the same way: replicate weights, a diagnostic check,
and an estimate. They differ in where the replicate step sits, and that
difference is the one rule worth memorizing.

### Probability sample

On a probability sample, create replicate weights **before** you
calibrate. `calibrate()` applied to a `survey_replicate` recalibrates
every replicate weight column to the same targets, so each replicate
carries its own copy of the adjustment. That per-replicate replay is
required: replicate weights built after the fact, from the already
calibrated full-sample weights, drop the variance the calibration step
contributes (Valliant, Dever & Kreuter 2018).

```{r probability-workflow}
# 1. Construct the design from the raw tibble. Calibration variables
#    must be NA-free, so drop the rows with NA sex or age first.
gss_complete <- gss_2024[!is.na(gss_2024$sex) & !is.na(gss_2024$age_f3), ]
gss <- surveycore::as_survey(
  gss_complete, weights = wtssps, strata = vstrat, ids = vpsu, nest = TRUE
)

# 2. Create replicate weights BEFORE calibrating.
#    Real analyses use more replicates; 100 keeps this article fast.
gss_rep <- create_replicate_weights(
  gss, method = "bootstrap", replicates = 100L, seed = 1
)

# 3. Calibrate to population margins — every replicate column is
#    recalibrated to the same targets
targets <- list(
  sex = c("Male" = 0.49, "Female" = 0.51),
  age_f3 = c("18-34" = 0.28, "35-54" = 0.37, "55+" = 0.35)
)
gss_cal <- calibrate(gss_rep, targets = targets)

# 4. Inspect the weights
summarize_weights(gss_cal)

# 5. Estimate
surveycore::get_freqs(gss_cal, pid_f3)
```

### Non-probability sample

A non-probability sample starts as a plain data frame. `ipw()` models
each unit's chance of appearing in the sample against a probability
reference survey and returns a `survey_nonprob`. Calibrating those
weights to census margins afterward gives a doubly robust estimate (see
the [glossary](#doubly-robust)).

Here the replicate step comes **last**. The quasi-randomization
bootstrap replays the whole weighting pipeline — the propensity model
and the calibration — inside every replicate, so the estimation error of
the models enters the variance. Calibrate the full-sample weights first;
create replicates as the final weighting step.

```{r nonprobability-workflow}
# 1. Start from a plain tibble; drop rows with NA in the benchmark
#    variable
ns_complete <- ns_wave1[!is.na(ns_wave1$pid_f3), ]

# 2. Estimate inverse probability weights against a probability
#    reference. With missing_method = "omit", ipw() warns and drops
#    rows with NA in a predictor.
npors_ref <- surveycore::as_survey(
  npors_2025_clean, weights = wt_pop, strata = stratum
)
nps_wts <- ipw(
  ns_complete,
  npors_ref,
  predictors = c("sex", "age_f3", "race_f4", "edu_f3"),
  missing_method = "omit",
  estimating_eq = "mle"
)

# 3. Calibrate the IPW weights to census margins (doubly robust)
nps_cal <- calibrate(nps_wts, targets = targets)

# 4. Create replicate weights LAST — each replicate refits the IPW
#    model and replays the calibration
nps_rep <- create_bootstrap_weights(
  nps_cal,
  type = "quasi-randomization",
  replicates = 100L,
  seed = 1
)

# 5. Inspect and estimate
summarize_weights(nps_rep)
surveycore::get_freqs(nps_rep, pid_f3)
```

The contrast in one sentence: on a probability sample the replicate
weights come first so calibration can be replayed onto them; on a
non-probability sample they come last because the replicate creator
replays the pipeline itself.

## Choosing a method

Each family's functions carry a "When to use" paragraph in their help
pages. The tables below give the short version.

### Calibration

`calibrate()` routes to the first three methods; `poststratify()` is a
separate function for full cell targets.

| Method | Choose it when | Watch for |
|---|---|---|
| `calibrate_rake()` (the default) | Your targets are separate margins — sex, age group, region — and the weights must stay positive (Deville, Sarndal & Sautory 1993). | The weight ratio has no upper bound. Use `cap`, or `trim_weights()` afterward. |
| `calibrate_linear()` | Speed matters: with `bounds = NULL` the solution is exact in one step, with no iteration (Deville, Sarndal & Sautory 1993). | Large gaps between sample and targets can produce negative weights. Set `bounds`, or switch method. |
| `calibrate_logit()` | You must limit how far any weight can move: the ratio of calibrated to starting weight stays inside `(L, U)` (Deville & Sarndal 1992). | Bounds that are too tight for the targets make a solution impossible, and the iteration fails. Widen `bounds` or rake. |
| `poststratify()` | You have population values for every joint cell of the stratification variables, and every cell contains sample members (Deville & Sarndal 1992). | With several variables the cells thin out quickly. When cells run empty, match margins via `calibrate()`. |

### Replicate weights

`create_replicate_weights()` routes to six methods. Each one assumes
something about how the sample was drawn.

| `method =` | Choose it when | Do not choose it when |
|---|---|---|
| `"bootstrap"` | Your estimand is a median or another quantile. Your estimator has no textbook variance formula. You have a non-probability sample plus a reference sample. | Your reference sample is stratified and multi-stage — that case is unproven. |
| `"jackknife"` (`"jkn"`) | You have a stratified, possibly multi-stage probability sample, every stratum holds 2 or more PSUs, and your estimand is a total, mean, or ratio. | Your estimand is a quantile. Any stratum holds one PSU. Every stratum holds exactly 2 PSUs — use `"brr"` instead. |
| `"jackknife"` (`"jk1"`) | You have a simple random sample with no strata and no clusters. | Same quantile limit as `"jkn"`. |
| `"jackknife"` (`"grouped"`) | You have a non-probability sample plus a reference sample, and you want fewer replicates than the ungrouped jackknife. Set `replicates` yourself. | Your estimand is a quantile. You need a proven consistency result for a doubly robust estimator — none exists. |
| `"brr"` | Every stratum holds exactly 2 PSUs, and especially when your estimand is a quantile or another nonlinear statistic. | Any stratum does not hold exactly 2 PSUs. Also set `rho > 0` if your estimand is a ratio. |
| `"generalized-bootstrap"` | Your design fits none of the above — Poisson sampling, an unusual multi-stage structure, or a case where you must name the target variance estimator yourself. | You cannot say which `variance_estimator` matches your design. Raise `replicates` to 750 or more. |
| `"generalized-replicate"` | You want the deterministic counterpart of generalized bootstrap, built from the same target variance estimators. | You planned to pass `replicates` — the argument does not exist. |
| `"successive-difference"` | Your sample was drawn systematically from a sorted list, and the rows still carry that sort order. | The rows are not in the original sort order. The sort variable carried no information — then the design was effectively a simple random sample. |

### Nonresponse

| Function and method | Choose it when |
|---|---|
| `adjust_nonresponse(method = "weighting-class")` | You can name the groups whose response rates differ, and those groups are observed for respondents and nonrespondents alike. |
| `adjust_nonresponse(method = "propensity-cell")` | No natural grouping exists: the model builds the cells from the data, and the cell-level factor smooths over individual fitted propensities. |
| `adjust_nonresponse(method = "propensity")` | Each unit's own fitted propensity should set its adjustment. Individual factors can move further than cell-level ones, so watch the `max_adjust` warning. |
| `redistribute_weights()` | The rows losing weight are defined by a condition other than nonresponse — for example, removing ineligible cases while conserving the group totals. |

### Calibrating to a survey vs. to estimates

| Function | Choose it when |
|---|---|
| `calibrate()` | Your targets are fixed census values with no sampling error of their own. |
| `calibrate_to_survey()` | The targets come from another survey and you have its data: the control survey's sampling error then carries into the replicate variance. |
| `calibrate_to_estimate()` | You have only published estimates and their variance-covariance matrix, not the control survey's data. |

## Glossary

### Design effect (DEFF)

How much the sampling design and the weights inflate the variance of an
estimate, compared with a simple random sample of the same size. A
design effect of 2 means the sample carries the information of an
unweighted sample half its size. `weight_variability()` reports the
weight CV, the quantity that drives the weighting component of the
design effect (Kish 1965).

### Doubly robust

An estimation strategy that combines two models — a selection model
(such as the IPW propensity model) and an outcome model (such as
calibration to population margins). The estimate stays consistent when
either model is right, so a mistake in one of the two is survivable. In
this package: `ipw()` followed by `calibrate()`.

### FPC (finite population correction)

A factor that shrinks a variance estimate when the sample is a large
share of the population. With a 5% sampling fraction it barely matters;
with 50% it halves the variance. Carried in the design object; not
something the weighting functions change.

### G-weight

The ratio of the calibrated weight to the starting weight,
`g = w_calibrated / w_start`. On the default multiplicative scale,
calibration `bounds` constrain this ratio rather than the weight itself:
`bounds = c(0.5, 2)` means no weight moves to less than half or more
than double its starting value. (`bounds_scale = "absolute"` switches to
bounding the weight directly.)

### Hadamard matrix

A square grid of +1 and -1 values arranged so that any two rows agree in
exactly half their positions. BRR, generalized replication, and
successive difference replication use one to keep the replicates
balanced — so that every unit is up-weighted and down-weighted in a
controlled pattern across replicates.

### MAR (missing at random)

The assumption that whether a unit responds depends only on things you
observed (age, region, mode), not on the unobserved answer itself. Every
nonresponse adjustment in this package leans on it, and no function can
test it for you.

### Propensity score

The modeled probability that a unit responds (in nonresponse adjustment)
or appears in the sample (in non-probability weighting). Weighting by
its inverse gives units from underrepresented groups more weight.

### PSU (primary sampling unit)

The first thing the design selects — often a county, a school, or a city
block, not a person. Replicate methods drop or reweight whole PSUs, not
individual respondents, because the PSU is the unit the randomness lives
in.

### Quasi-randomization bootstrap

The bootstrap for a non-probability sample with a probability reference
sample. Each replicate resamples the non-probability sample and refits
the weighting model, so the model's own estimation error enters the
variance; the reference sample is also resampled when the calibration
targets were estimated from it. See
`create_bootstrap_weights(type = "quasi-randomization")`.

### Replicate weights

Sets of perturbed copies of the main weight, stored as extra columns.
You compute your estimate once per column; the spread of those estimates
gives the standard error. They exist so that variance estimation
survives complex designs and multi-step weighting.

### Taylor linearization

The closed-form alternative to replication: approximate a nonlinear
statistic by a linear one, then apply the design's exact variance
formula. Fast and standard for totals, means, and ratios; it is what a
`survey_taylor` object uses.

### Weighting class

A group of units assumed to share the same response rate.
`adjust_nonresponse(method = "weighting-class")` moves the weight of
nonrespondents to respondents inside each class.

### Method names

- **BRR** — balanced repeated replication; needs exactly two PSUs per
  stratum. See `create_brr_weights()`.
- **DAGJK** — delete-a-group jackknife; the grouped jackknife for
  non-probability samples. See
  `create_jackknife_weights(type = "grouped")`.
- **GREG** — generalized regression estimator; the survey name for
  linear calibration. See `calibrate_linear()`.

## Further reading

Deville, J.-C. and Sarndal, C.-E. (1992). Calibration estimators in
survey sampling. *Journal of the American Statistical Association*,
87(418), 376--382.

Deville, J.-C., Sarndal, C.-E. and Sautory, O. (1993). Generalized
raking procedures in survey sampling. *Journal of the American
Statistical Association*, 88(423), 1013--1020.

Kish, L. (1965). *Survey Sampling*. New York: John Wiley & Sons.

Valliant, R., Dever, J. and Kreuter, F. (2018). *Practical Tools for
Designing and Weighting Survey Samples*, 2nd edition. New York:
Springer.
